---
title: "Interactive causal diagrams"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive causal diagrams}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A directed acyclic graph (DAG) sets out the causal assumptions behind an
analysis: which variables are included, and which way each effect is thought
to run. A review of DAGs in applied health research found wide variation in
how they are reported and recommended making them more transparent
([Tennant et al. 2021](https://doi.org/10.1093/ije/dyaa213)). A printed
diagram has little room for the reasoning behind each node and arrow.
`ggcausal()` keeps that reasoning with the diagram. Every node and every
arrow carries its own rationale and references, shown when the reader hovers
over it and opened in full, with clickable links, when they click.

```{r setup}
library(ggextreme)
```

## A first diagram

The package includes a small illustrative DAG for a study of maternal smoking
and orofacial clefts in the child.

```{r}
dag <- ggcausal(
  cleft_dag$edges,
  cleft_dag$nodes,
  legend_title = "Role",
  title = "Maternal smoking and orofacial clefts"
)
dag
```

Hover over a box or an arrow to see why it is there. Click it to open the
full rationale and the references under the diagram; click it again, or the
close button, to hide them. Clicking works on touch screens, where there is
no hover.

## The data

A diagram is two data frames. `edges` has one row per arrow and is the only
required input:

```{r}
cleft_dag$edges[c("from", "to")]
```

`nodes` has one row per variable, with a `name` matching `from` and `to`:

```{r}
cleft_dag$nodes[c("name", "label", "role")]
```

A few column names have a meaning of their own.

| column | in | effect |
| --- | --- | --- |
| `from`, `to` | edges | the ends of the arrow |
| `name` | nodes | the identifier used in `from` and `to` |
| `label` | nodes | text in the box, which may contain line breaks |
| `role` | nodes | sets the color, and appears in the legend |
| `rationale` | both | why the node or arrow is in the diagram |
| `references` | both | sources for the rationale |
| `x`, `y` | nodes | place the box by hand |

Every other column is shown as a labeled field. The example has a `timing`
column on its nodes, which appears as "Timing" in the hover card and the
panel. This is the place for anything a reader might want to check, such as
how a variable was measured.

## References

`references` holds one or more references per row. Give several either as a
list column or as text separated by `|` or a line break. Semicolons are left
alone because Vancouver style uses them inside a single reference.

```{r}
cleft_dag$nodes$references[2]
```

A reference that contains a URL is linked to it. Otherwise one that contains
a DOI, with or without a `doi:` prefix, is linked to `https://doi.org/`.
All text is escaped, so characters such as `<` and `&` appear as written.

## Roles

`role` is free text. Eight roles have fixed colors: `exposure`, `outcome`,
`confounder`, `mediator`, `collider`, `instrument`, `unobserved` and
`latent`. The last two are drawn with a dashed border. Any other role, such
as the `protective factor` in the example, takes the next color from
`race_palette()`. Override any of them with a named `palette`:

```{r}
ggcausal(
  cleft_dag$edges, cleft_dag$nodes,
  palette = c(exposure = "#1B6CA8", outcome = "#B8413A")
)
```

## Layout

By default the nodes are placed in layers so that every arrow points the same
way, left to right or, with `direction = "down"`, top down. An arrow that
skips a layer is routed around the boxes in between on a smooth curve.

```{r}
ggcausal(cleft_dag$edges, cleft_dag$nodes, direction = "down", legend = FALSE)
```

For full control, give `x` and `y` columns in `nodes`. They are grid
positions with `y` pointing up, and the arrows between them are straight.

```{r}
nodes <- data.frame(
  name = c("confounder", "exposure", "outcome"),
  role = c("confounder", "exposure", "outcome"),
  x = c(1, 0, 2),
  y = c(1, 0, 0)
)
edges <- data.frame(
  from = c("confounder", "confounder", "exposure"),
  to = c("exposure", "outcome", "outcome")
)
ggcausal(edges, nodes)
```

## Other forms

The diagram prints as a widget in the RStudio viewer, in R Markdown and
Quarto documents, and on 'pkgdown' sites. Three functions give its other
forms:

* `graph_widget()` returns the htmlwidget, for 'shiny' or
  `htmlwidgets::saveWidget()`.
* `graph_plot()` returns the underlying ggplot, a static diagram for a
  manuscript.
* `graph_save()` writes the widget to a single `.html` file, which can go in
  a supplement, or a static `.png` at its natural size.

```{r, eval = FALSE}
graph_save(dag, "dag.html")
graph_save(dag, "dag.png", res = 300)
```

The diagram is laid out at a fixed size so that every box fits its label.
The widget scales down to fit a narrow page; a static copy should be drawn at
`dag$width` by `dag$height` inches, which is what `graph_save()` does.
