---
title: "Confidence in a network meta-analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Confidence in a network meta-analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = requireNamespace("netmeta", quietly = TRUE) &&
    requireNamespace("meta", quietly = TRUE)
)
```

A network meta-analysis gives an estimate for every pair of treatments, but
not every estimate deserves the same confidence. CINeMA, Confidence in
Network Meta-Analysis (Nikolakopoulou et al. 2020; Papakonstantinou et al.
2020), judges each estimate in six domains: within-study bias, reporting
bias, indirectness, imprecision, heterogeneity and incoherence. The
`cinema_` functions apply its published rules and draw what lies behind each
judgment, so a reader can see why an estimate earns the confidence it gets:

* `cinema_judge()` gives every comparison a judgment in each domain, with
  the reason in words and whether a rule computed it or you gave it;
* `cinema_contribution()` shows which studies each estimate rests on;
* `cinema_clinical()` sets the estimates against a range of little
  difference that the reader can move;
* `cinema_incoherence()` puts direct, indirect and network estimates side
  by side;
* `cinema_league()` draws the league table with the six judgments in every
  cell;
* `cinema_network()` draws the network with one strand per study, colored
  by its judgment.

```{r setup}
library(ggextreme)
```

## The network and some illustrative judgments

The network is the package's five trials of treatments for plaque psoriasis,
analyzed for PASI 75 response as odds ratios with a random effects model.

```{r}
pw <- meta::pairwise(
  treat = treatment, event = pasi75_r, n = pasi75_n,
  studlab = study, data = psoriasis_nma, sm = "OR"
)
nma <- netmeta::netmeta(pw, common = FALSE)
```

CINeMA starts from a judgment of each study's risk of bias and of its
indirectness, its relevance to the review question. **The judgments below
are illustrative, invented for this vignette; they are not published
assessments of these trials.** Each is a data frame with the study, the
judgment and, optionally, the reason for it.

```{r}
rob <- data.frame(
  study = c("CLEAR", "ERASURE", "FEATURE", "FIXTURE", "JUNCTURE"),
  judgment = c("high", "low", "some concerns", "low", "some concerns"),
  reason = c(
    "Illustrative: the reported PASI 75 analysis could not be matched to a prespecified plan.",
    "Illustrative: placebo controlled and double blind, with almost everyone analyzed.",
    "Illustrative: the masking of outcome assessors was poorly reported.",
    "Illustrative: double blind against placebo and etanercept, with almost everyone analyzed.",
    "Illustrative: allocation concealment was poorly reported."
  )
)
indirectness <- data.frame(
  study = rob$study,
  judgment = c("low", "low", "high", "low", "moderate"),
  reason = c(
    "Illustrative: population, doses and outcome match the review question.",
    "Illustrative: population, doses and outcome match the review question.",
    "Illustrative: participants had to be willing to self-inject, and BMI was not reported.",
    "Illustrative: population, doses and outcome match the review question.",
    "Illustrative: an autoinjector trial that may favor people comfortable with self-injection."
  )
)
```

Reporting bias cannot be computed from the data, so it is your judgment:
undetected or suspected, for every comparison at once or one by one.

```{r}
reporting <- data.frame(
  judgment = "suspected",
  reason = "Illustrative: every trial was funded by the maker of one of its drugs."
)
```

## Judging every comparison

`cinema_judge()` applies CINeMA's rules. Imprecision, heterogeneity and, in
part, incoherence need a range of little difference: the effects too small
to matter to patients, chosen before looking at the results. Here it is an
odds ratio from 0.8 to 1.25, given as one number. For a response, larger
odds ratios are better, so `small_values = "undesirable"`.

```{r}
j <- cinema_judge(
  nma, rob = rob, indirectness = indirectness, reporting = reporting,
  threshold = 1.25, small_values = "undesirable"
)
j
```

Each judgment comes with its reason and its source:

```{r}
imp <- j$judgments[j$judgments$domain == "Imprecision", ]
imp$reason[5]
```

Computing the contribution of each study, with `netmeta::netcontrib()`,
takes a few seconds, so judge once and give the result to each plot. Every
plot also accepts the netmeta fit with the arguments of `cinema_judge()`.

## Where each estimate's evidence comes from

An estimate's within-study bias and indirectness depend on the studies it
rests on, weighted by how much each contributes. `cinema_contribution()`
splits each estimate into its studies, grouped low, moderate and high as
CINeMA draws them, beside a comparison by study matrix and a small network.

```{r}
cinema_contribution(j, caption = "Illustrative judgments, not published assessments.")
```

Switch between risk of bias and indirectness; select a comparison, a study
or part of a bar, and the bars, the matrix and the network follow. Selecting
Ustekinumab vs Placebo shows that no trial compares them directly, and that
two fifths of the estimate flows through CLEAR, the trial with the high
illustrative risk of bias.

## Against a range of little difference

`cinema_clinical()` draws each network estimate with its confidence and
prediction intervals against the range of little difference, and says in
words what each interval is compatible with. The sliders move the lower and
upper limits independently, while the prespecified limits stay marked; the
readings, the imprecision and heterogeneity judgments at those limits and
the sensitivity strips under the estimates follow.

```{r}
cinema_clinical(j)
```

The strips show, for each comparison, where the reading changes as one
limit moves with the other held. Here Secukinumab 300 mg vs Secukinumab 150
mg would also be compatible with little difference if the upper limit were
raised above 1.33, the lower end of its interval, and the two comparisons
with ustekinumab whose intervals start near 0.95 would be compatible with an
important harm if the lower limit were raised past it. With five trials the
between-study variance is poorly estimated, so the prediction intervals are
only a rough guide.

## Direct and indirect evidence

`cinema_incoherence()` places the direct estimate, from the trials of each
pair, beside the indirect estimate from the rest of the network, with the
inconsistency factor and its confidence interval. Comparisons with only
direct or only indirect evidence cannot be checked locally and are marked as
such; CINeMA judges them from the global design by treatment test instead.

```{r}
cinema_incoherence(j)
```

Wide intervals of the inconsistency factor, such as those of the two
secukinumab doses against placebo, show how weak these tests are: a large
p-value does not mean that direct and indirect evidence agree.

## The confidence profile of every estimate

`cinema_league()` lays out the league table as `ggleague()` does, network
estimates below the diagonal and direct estimates above it, and adds six
marks under each network estimate, one per domain. Click a cell for every
judgment with its reason and source, and a bar of the estimate's
contributions.

```{r}
cinema_league(j, caption = "Illustrative judgments, not published assessments.")
```

The marks are never added into a score. CINeMA leaves any overall rating to
the reviewers, and warns that the domains are related: one trial at high
risk of bias can raise concerns in more than one of them.

## Judgments on the network

`cinema_network()` draws the network from the arm level data with each line
split into one strand per study, so a comparison that mixes studies at low
and high risk of bias shows both colors instead of an average. It takes the
study judgments directly; here the treatments are placed by hand so that no
lines cross.

```{r}
positions <- data.frame(
  treatment = c("Secukinumab 300 mg", "Ustekinumab", "Secukinumab 150 mg",
                "Etanercept", "Placebo"),
  x = c(210, 372, 62, 358, 210),
  y = -c(40, 40, 200, 200, 138)
)
cinema_network(
  psoriasis_nma, study, treatment, n = n,
  rob = rob, indirectness = indirectness, positions = positions,
  caption = "Illustrative judgments, not published assessments."
)
```

## Judgments you made yourself

A judgment the rules would get wrong can be replaced, and judgments made
elsewhere, such as the report the CINeMA web application exports, can be
drawn as they are. Give them in `judgments`, wide with one column per domain
or long with a `domain` column; they are labeled as yours in every plot.

```{r}
mine <- data.frame(
  comparison = "Secukinumab 150 mg:Ustekinumab",
  Imprecision = "major concerns",
  "Imprecision reason" = "Illustrative: the upper limit we would accept is lower.",
  check.names = FALSE
)
j2 <- cinema_judge(
  nma, rob = rob, reporting = reporting, threshold = 1.25,
  small_values = "undesirable", judgments = mine
)
j2$judgments[j2$judgments$domain == "Imprecision", c("treat1", "treat2", "judgment", "source")][5, ]
```

## The rules

| Domain | Rule | Source of the judgment |
| --- | --- | --- |
| Within-study bias | study judgments weighted by their contributions, by the average, majority or highest rule | computed from your study judgments |
| Reporting bias | none: undetected or suspected | yours |
| Indirectness | as within-study bias | computed from your study judgments |
| Imprecision | the CI against the range of little difference and the side of no effect | computed |
| Heterogeneity | the prediction interval by the same rule, compared with the CI | computed |
| Incoherence | the local test of direct against indirect evidence, and where it is 0.10 or less the areas both CIs reach; otherwise the global test | computed |

The help page of `cinema_judge()` gives each rule in full, with the choices
made where the papers leave a detail open.

## References

Nikolakopoulou A, Higgins JPT, Papakonstantinou T, et al. CINeMA: an
approach for assessing confidence in the results of a network
meta-analysis. PLoS Medicine 2020;17(4):e1003082.

Papakonstantinou T, Nikolakopoulou A, Higgins JPT, Egger M, Salanti G.
CINeMA: software for semiautomated assessment of the confidence in the
results of network meta-analysis. Campbell Systematic Reviews
2020;16:e1080.

Papakonstantinou T, Nikolakopoulou A, Rucker G, et al. Estimating the
contribution of studies in network meta-analysis: paths, flows and streams.
F1000Research 2018;7:610.
