---
title: "Funnel plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Funnel plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = requireNamespace("metafor", quietly = TRUE) &&
    requireNamespace("metadat", quietly = TRUE)
)
```

A funnel plot asks whether the small studies in a meta-analysis tell a
different story from the large ones. `ggfunnel()` draws it from a fitted
model, shades where each study would be significant against no effect, and
keeps the evidence behind every point a hover or a click away: the study's
record, the pooled estimate without it, and the tests for small-study
effects.

```{r setup}
library(ggextreme)
```

## A first funnel plot

The example is the classic set of 13 trials of the BCG vaccine against
tuberculosis, which ships with the metadat package, fitted as a random
effects model of the risk ratio.

```{r}
dat <- metafor::escalc(measure = "RR", ai = tpos, bi = tneg, ci = cpos, di = cneg,
                       data = metadat::dat.bcg, slab = paste(author, year))
fit <- metafor::rma(yi, vi, data = dat)

f <- ggfunnel(fit, hover = c("alloc", "ablat"), trim_fill = TRUE,
              title = "BCG vaccine and tuberculosis")
f
```

Each point is a study, sized by its weight, with the most precise at the
top. The solid line is the pooled estimate and the dashed lines around it
the region where 95% of studies would fall without heterogeneity or bias.
Hover over a study for its risk ratio, weight, the columns named in `hover`
and the zone it sits in; click it to compare the pooled estimate with and
without it.

## Reading the contours

The shading is centered on no effect, not on the pooled estimate (Peters et
al. 2008). A study in a shaded band would reach that level of significance
on its own; one in the unshaded middle would not. If studies are missing
from a funnel only where they would not have been significant, publication
bias is a likely cause. If they are missing from shaded areas too, the
asymmetry more likely comes from somewhere else, such as heterogeneity or
the conduct of the small studies. `contours` sets the levels, and
`contours = NULL` leaves the shading out.

## Tests for small-study effects

The section under the plot, collapsed until opened, holds Egger's regression
test, Begg's rank correlation test and, with `trim_fill = TRUE`, the trim and
fill estimate, as computed by metafor, or by meta for a meta object.
Egger's test is the classical one, `metafor::regtest(fit, model = "lm")`,
which meta also computes; metafor's default `regtest(fit)` fits a
meta-regression instead and gives a different p value. The two packages'
trim and fill can impute different numbers of studies from the same data:
on these trials metafor imputes one and meta four. The table is also
returned:

```{r}
f$tests
```

Here neither test finds asymmetry, and trim and fill imputes one study on
the right, which moves the risk ratio little. The imputed study is drawn as
a hollow circle and the adjusted estimate as a dashed line, and the switch
under the widget hides them. These tests have little power with fewer than
ten studies, and asymmetry is not proof of publication bias.

## Risk of bias

`rob` names the column holding each study's overall judgement, which colors
its point, with the wording matched as in `ggmeta()`:

```{r}
chi <- metafor::escalc(measure = "OR", ai = p2y12.mi, n1i = p2y12.total,
                       ci = aspirin.mi, n2i = aspirin.total,
                       data = metadat::dat.chiarito2020, slab = paste(study, year))
chi <- chi[!is.na(chi$yi), ]
ggfunnel(metafor::rma(yi, vi, data = chi), rob = "rob.overall",
         title = "P2Y12 inhibitors against aspirin: myocardial infarction")
```

## Options

| argument | effect |
| --- | --- |
| `x` | an `rma.uni` fit from metafor, or a `meta` object |
| `hover` | columns of the data for each study's hover card |
| `rob` | the column of overall risk of bias judgements |
| `contours` | significance levels for the shading, or `NULL` |
| `trim_fill` | add the imputed studies and the adjusted estimate |
| `tests` | add the collapsed section of tests |
| `exponentiate`, `xlim`, `xlab` | the scale, limits and label of the effect axis |

On a dark page the plot takes a dark palette of its own, and
`graph_save(f, "funnel.png", theme = "dark")` writes a dark static copy.
