---
title: "A multiverse of analyses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A multiverse of analyses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Any analysis rests on choices: how the outcome is defined, what is adjusted
for, which model is fitted, how missing data are handled, who is included.
When several choices are defensible, running every combination shows how
much the answer depends on them. `ggmultiverse()` draws the result as a
specification curve and shows which choices move it.

```{r setup}
library(ggextreme)
```

## A first multiverse

The analyses below are simulated: every combination of five choices, 48 in
all, each with an odds ratio and its interval.

```{r}
specs <- expand.grid(
  outcome = c("Primary definition", "Broad definition"),
  adjustment = c("Minimal", "Standard", "Extended"),
  model = c("Logistic", "Log-binomial"),
  missing = c("Complete cases", "Multiple imputation"),
  sample = c("Everyone", "Excluding early deaths"),
  stringsAsFactors = FALSE
)
set.seed(7)
b <- -0.35 + 0.15 * (specs$outcome == "Broad definition") +
  c(0, 0.12, 0.2)[match(specs$adjustment, c("Minimal", "Standard", "Extended"))] -
  0.03 * (specs$model == "Log-binomial") + 0.05 * (specs$missing == "Multiple imputation") -
  0.08 * (specs$sample != "Everyone") + rnorm(48, 0, 0.03)
specs$or <- exp(b)
specs$lo <- exp(b - 1.96 * 0.09)
specs$hi <- exp(b + 1.96 * 0.09)

mv <- ggmultiverse(
  specs, or, lo, hi,
  decisions = c("outcome", "adjustment", "model", "missing", "sample"),
  primary = outcome == "Primary definition" & adjustment == "Standard" &
    model == "Logistic" & missing == "Multiple imputation" & sample == "Everyone",
  ylab = "Odds ratio", title = "Forty-eight ways to analyze one question"
)
mv
```

Each column is one analysis: its estimate and interval above, and below, a
dot in every row of the grid for the choice it made. Colors say whether the
interval lies below 1, above it, or includes it. The primary analysis is
labeled. Beside each choice sits the median estimate of the analyses that
made it, so the choices that matter stand out: here the adjustment set moves
the median from 0.75 to 0.90.

Drag across the curve to select a run of analyses, and the readout says
which choices they share. Click a choice in the grid to keep only the
analyses that made it; clicks on several choices combine, and the primary
analysis stays in view.

```{r}
mv$influence
```

## Reading it well

The share of analyses with intervals that exclude 1 describes this set of
analyses. It is not a probability that the effect is real, and the analyses
are far from independent. More than 500 analyses cannot be read as a curve,
so `ggmultiverse()` asks for fewer.
