---
title: "Interactive network plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interactive network plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A network meta-analysis compares several treatments at once by combining
trials that each compare a few of them. The network plot shows which
comparisons the evidence covers: every treatment is a node, and every pair
of treatments compared directly in at least one trial is joined by a line.
Judging whether the trials are similar enough to combine means going back to
the arm level data, the baseline characteristics and outcomes of each arm.
`ggnma()` puts that data behind the plot. Hovering over a node or a line
shows its arms side by side, in the manner of a trial's baseline table, and
clicking it opens the full table with every column.

```{r setup}
library(ggextreme)
```

## A first network

The package includes arm level data from five randomized trials in moderate
to severe plaque psoriasis, compiled by Phillippo et al. for their
multilevel network meta-regression.

```{r}
net <- ggnma(
  psoriasis_nma,
  study = study,
  treatment = treatment,
  n = n,
  group = class,
  legend_title = "Class",
  title = "Treatments for plaque psoriasis"
)
net
```

Hover over the line between placebo and secukinumab 150 mg to compare the
four trials that make that comparison, arm by arm, and click it for every
column. Hover over or click a node to see every arm on that treatment. The
references at the bottom of each panel are linked.

## The data

`ggnma()` takes long data with one row per study arm, the shape
`netmeta::pairwise()` and `multinma::set_agd_arm()` start from:

```{r}
head(psoriasis_nma[c("study", "treatment", "n", "pasi75_r", "age")])
```

Three columns place each arm: `study`, `treatment` and, optionally, `group`
for the drug class. Every other column becomes a row of the panel tables,
in the order given, so choosing what to show is a matter of selecting
columns first. `n` is optional; when given, node area follows the total
number of participants on each treatment, and the hover cards count
participants as well as studies.

Rows are named from each column's `label` attribute when it has one, as set
by the 'labelled', 'Hmisc' or 'haven' packages, and otherwise from the
column name:

```{r}
attr(psoriasis_nma$age, "label")
```

Text that is the same for every arm of a study, such as the reference here,
is listed once per study under the table rather than repeated in every
column. Text that contains a URL or a DOI is linked.

## The hover card

The hover card holds a compact version of the table. By default it shows
the first four columns; `hover` picks them by name, so the card can carry
the sample size, the outcome and the baseline characteristic that matters
most for the question at hand:

```{r}
ggnma(
  psoriasis_nma, study, treatment, n = n, group = class,
  hover = c("n", "pasi75_r", "pasi_w0", "prior_systemic"),
  legend = FALSE
)
```

`hover = character(0)` gives a card that only lists the studies. The click
panel always shows every column.

## What the plot encodes

* **Nodes** are treatments, placed on a circle starting at the top and
  running clockwise in the order of the factor levels of `treatment`, or in
  order of first appearance. `positions` places them by hand.
* **Node area** follows the number of participants on each treatment when
  `n` is given.
* **Line width** follows the number of studies that make the comparison.
* **A multi-arm study** adds a line for every pair of its treatments, and a
  shaded polygon joining them. The counts are available on the object:

```{r}
net$edges
```

## Multi-arm studies

Lines alone cannot show that three treatments were compared within one
trial rather than in three separate ones. A shaded polygon joins the
treatments of every study with more than two arms, and studies that compare
the same set share one polygon. Here ERASURE, FEATURE and JUNCTURE each
compare placebo with both doses of secukinumab, and FIXTURE adds etanercept:

```{r}
net$multiarm
```

Each polygon has its own hover card and panel, with the arms of every study
it stands for. `multiarm = FALSE` leaves the shading out.

## Placing nodes by hand

`positions` takes one row per treatment with `x` and `y` coordinates, `y`
pointing up. Any units will do: the layout is scaled to fit the plot,
keeping its shape, and labels point away from its middle. This one puts
placebo at the top, the active comparators at the sides and the two
secukinumab doses along the bottom:

```{r}
positions <- data.frame(
  treatment = c("Placebo", "Etanercept", "Ustekinumab",
                "Secukinumab 150 mg", "Secukinumab 300 mg"),
  x = c(1, 0, 2, 0.5, 1.5),
  y = c(1, 0, 0, -1, -1)
)
ggnma(psoriasis_nma, study, treatment, n = n, group = class,
      positions = positions, legend = FALSE)
```

## Colors

Without `group`, every node takes the first color of `race_palette()`, or
the single color in `palette`. With `group`, nodes are colored by class and
a legend is drawn. A named `palette` sets colors by class, and classes left
out keep their default:

```{r}
ggnma(
  psoriasis_nma, study, treatment, n = n, group = class,
  palette = c(Placebo = "#9A9A9A"),
  legend_title = "Class"
)
```

## Where the evidence for a comparison comes from

Give a netmeta fit on the same network, or the result of
`netmeta::netcontrib()` for it, as `contributions`, and the widget gains a
menu of every comparison. Picking one widens and colors each line by the
share of that network estimate flowing through it, labels the shares and
lists them under the plot, while the lines it does not use fade.

```{r, eval = requireNamespace("netmeta", quietly = TRUE) && requireNamespace("meta", quietly = TRUE)}
pw <- meta::pairwise(treat = treatment, event = pasi75_r, n = pasi75_n,
                     studlab = study, data = psoriasis_nma, sm = "OR")
fit <- netmeta::netmeta(pw, common = FALSE)
ggnma(psoriasis_nma, study, treatment, n = n, group = class,
      legend_title = "Class", contributions = fit)
```

No trial compares ustekinumab with placebo, so its estimate flows entirely
through the other comparisons, most of it through the one trial of
ustekinumab against secukinumab.

## Other forms

As with causal diagrams, the plot prints as a widget in the RStudio viewer,
in R Markdown and Quarto documents, and on 'pkgdown' sites.
`graph_widget()` returns the htmlwidget for 'shiny', `graph_plot()` the
underlying ggplot for a manuscript, and `graph_save()` a single `.html` file
or a static `.png`:

```{r, eval = FALSE}
graph_save(net, "network.html")
graph_save(net, "network.png", res = 300)
```

## Source of the example data

The arm level data are those compiled by Phillippo (2019) and distributed
with the 'multinma' package, from the published reports of CLEAR, ERASURE,
FEATURE, FIXTURE and JUNCTURE; see `?psoriasis_nma`. They were analyzed in
[Phillippo et al. (2020)](https://doi.org/10.1111/rssa.12579).
