---
title: "Responder thresholds"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Responder thresholds}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Trials often report the share of patients who improved by at least some
amount, the responders. The threshold makes the result easy to state, and
throws away how much each patient changed. `ggresponder()` keeps both in
view: the whole distribution of change, the responders at the prespecified
threshold, and how the difference would look at any other.

```{r setup}
library(ggextreme)
```

## A first plot

In this simulated trial, pain is scored from 0 to 10 and a drop of 2 points
is the prespecified threshold for a responder. `higher_is_better = FALSE`
turns the change around so that improvement is positive:

```{r}
set.seed(3)
pain <- data.frame(
  arm = factor(rep(c("Placebo", "Active"), each = 120), levels = c("Placebo", "Active")),
  change = c(rnorm(120, -1.3, 2), rnorm(120, -2.2, 2))
)
r <- ggresponder(change ~ arm, pain, threshold = 2, higher_is_better = FALSE,
                 xlab = "Improvement in pain (points on a 0 to 10 scale)",
                 title = "Who counts as a responder?")
r
```

On the left, each curve gives the share of an arm who improved by at least
each amount; the threshold picks one point on each. On the right, the
difference in responders is drawn at every threshold with its confidence
band: its shape shows how much the conclusion depends on where the line is
drawn. Under them, the table gives the responders in each arm, their
difference, the number needed to treat and the mean difference, which uses
every patient.

Move the slider, or drag either panel, to try another threshold; a button
returns to the prespecified one. A threshold picked after seeing the data
can make almost any difference look real, and the readout says so.

```{r}
r$responders
```

## Intervals

Responders have Wilson score intervals and their difference the hybrid
score interval of Newcombe. When the interval of the difference includes
zero, the interval of the number needed to treat runs from benefit through
infinity to harm, as Altman describes. The mean difference has a Welch
interval.
