---
title: "Bias and tipping points"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bias and tipping points}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

Every observational estimate invites the same question: how strong would
unmeasured confounding have to be to change the conclusion? `ggsensitivity()`
answers it for every combination of strengths at once, and marks where the
answer changes.

```{r setup}
library(ggextreme)
```

## A first explorer

Suppose an analysis gives a risk ratio of 1.80 (1.40 to 2.31), that a risk
ratio of 1.25 is the smallest that would matter clinically, and that the two
strongest measured confounders, age and smoking, were associated with the
exposure and the outcome as below:

```{r}
bench <- data.frame(label = c("Age", "Smoking"),
                    exposure = c(1.6, 2.3), outcome = c(1.9, 1.5))
s <- ggsensitivity(1.8, 1.4, 2.31, important = 1.25, benchmarks = bench,
                   title = "How strong would confounding have to be?")
s
```

Each point of the surface is an unmeasured confounder with a risk ratio
with the exposure on one axis and with the outcome on the other. The shading
says what would survive it: a clinically important effect with an interval
clear of 1, an interval clear of 1, an estimate on the same side of 1, or
nothing. Click the surface, or use the sliders, to choose a confounder; the
row "Under the chosen confounder" and the sentence under the plot follow.

```{r}
s$evalues
s$benchmarks
```

The E-value is the strength, on both axes at once, that could just explain
the estimate away; the second E-value does the same for the confidence
limit. A confounder as strong as the benchmarks would leave the estimate
clear of 1, which is some reassurance, not proof: an unmeasured confounder
need not resemble the measured ones.

## The method

The adjustment divides the estimate by the bounding factor of Ding and
VanderWeele, which is the most bias a confounder of given strengths could
produce, so every adjusted value is a worst case. A protective estimate is
handled by symmetry. An odds ratio or hazard ratio for a common outcome is
first converted to an approximate risk ratio; set `rare = TRUE` when the
outcome is rare.

```{r}
ggsensitivity(0.62, 0.48, 0.80, measure = "HR", important = 0.8)
```
