---
title: "Swimmer plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Swimmer plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = requireNamespace("survival", quietly = TRUE)
)
```

A swimmer plot gives every patient a lane: how long they were on treatment
or on study, when they responded, when they progressed and who is still
going. It is the standard picture of response in single arm and early phase
trials, where the course of each patient matters as much as the summary.
`ggswimmer()` draws one with each lane's record under the pointer.

```{r setup}
library(ggextreme)
```

## A first swimmer plot

The survival package ships `myeloid`, a simulated data set based on a trial
in acute myeloid leukemia, with the times of complete response, stem cell
transplant, relapse and death. The first 30 patients make a readable plot.
The events go in a data frame of their own, one row per event, with the
patient's id, the time and what happened:

```{r}
aml <- subset(survival::myeloid, id <= 30)
aml$arm <- paste("Arm", aml$trt)
month <- 30.44
events <- rbind(
  data.frame(id = aml$id, time = aml$crtime / month, event = "Complete response"),
  data.frame(id = aml$id, time = aml$txtime / month, event = "Transplant"),
  data.frame(id = aml$id, time = aml$rltime / month, event = "Relapse"),
  data.frame(id = aml$id, time = ifelse(aml$death == 1, aml$futime / month, NA),
             event = "Death")
)
events <- events[!is.na(events$time), ]

ggswimmer(aml, id, futime / month, events = events, group = arm,
          ongoing = death == 0, hover = c("sex", "flt3"),
          ongoing_label = "Alive at last follow-up",
          xlab = "Months since randomization",
          title = "Acute myeloid leukemia, 30 patients")
```

Hover over a lane to read the patient's arm, the columns named in `hover`,
the first time of each event and whether they are still being followed;
the other lanes fade. Click it for the patient's events in order, beside
their full record. The buttons under the plot reorder the lanes, longest
first, by arm, by best response or in the order of the data, and the lanes
slide into place.

## Events

`end` and `ongoing` take expressions of the columns of `data`, as above.
Events are drawn by their wording, whatever the case:

| wording | mark |
| --- | --- |
| complete, CR | star |
| partial, PR, response, remission | triangle |
| progression, PD, relapse, recurrence | diamond |
| death, died | cross |
| anything else | a circle or square in its own color |

## Waterfall and trajectories

In a solid tumor trial the lanes are only half the story: how far each
tumor shrank matters too. `waterfall` gives each patient's best percent
change from baseline, drawn beside their lane, and `trajectories` their
change at each assessment, drawn under the lanes on the same time axis. The
three panels share one row, or one line, per patient: hovering over any of
them lights the patient in all three. The order "Best change" sorts the lanes
as a waterfall, and the response and progression thresholds, by default the
30 percent decrease and 20 percent increase of RECIST 1.1 for target
lesions, are marked on both change panels with a count past each.

```{r}
set.seed(4)
n <- 16
pts <- data.frame(patient = sprintf("P%02d", 1:n), arm = rep(c("Drug", "Control"), each = n / 2),
                  months = round(pmax(2, rexp(n, 1 / 10)), 1))
slope <- ifelse(pts$arm == "Drug", -9, -2) + rnorm(n, 0, 5)
traj <- do.call(rbind, lapply(1:n, function(i) {
  t <- seq(2, pts$months[i], by = 2)
  if (!length(t)) return(NULL)
  data.frame(patient = pts$patient[i], time = t,
             change = round(pmax(-100, cumsum(slope[i] + rnorm(length(t), 0, 6)))))
}))
pts$best <- as.numeric(tapply(traj$change, traj$patient, min)[pts$patient])

ggswimmer(pts, patient, months, group = arm, waterfall = best,
          trajectories = traj, sort = "change", xlab = "Months since first dose")
```

These data are simulated. The measured change is not the whole of RECIST:
new lesions and other progression count too. `response` in the result holds
the counts past each threshold.

## Options

| argument | effect |
| --- | --- |
| `start` | where each bar starts; zero by default |
| `group` | the column that colors the bars, such as the arm |
| `ongoing`, `ongoing_label` | who gets an arrow, and what it means |
| `sort` | the order of a static copy: `"duration"`, `"group"`, `"response"`, `"change"` or `"data"` |
| `palette` | colors for the groups |
| `hover` | columns for each lane's hover card |
| `waterfall`, `trajectories` | best change beside the lanes, and change over time under them |
| `thresholds`, `change_label` | the response and progression thresholds, and the axis label |

On a dark page the plot takes a dark palette of its own, and
`graph_save(s, "swimmer.png", theme = "dark")` writes a dark static copy.
